---
title: "RWorksheet_orada#4a.Rmd"
author: "Jessamine Paula Orada"
date: "2025-11-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#1. The table below shows the data about shoe size and height. Create a data frame.
#a. Describe the data.
##The data, presented in Figure 1, is a collection of measurements related to shoe size and height from various respondents, along with their Gender.Variables:

#Gender: A nominal categorical variable (or factor) with levels 'M' (Male) and 'F' (Female).

#Shoe Size: A quantitative (discrete) variable, likely measured in standard shoe size units.

#Height: A quantitative (continuous) variable, likely measured in centimeters or inches.

#Data Type: This dataset is suitable for exploratory data analysis to examine the relationship between shoe size and height, and to compare these measurements across genders.



#1.b. Create a subset by males and females with their corresponding shoe size and height.What its result? Show the R scripts.

```{r}
# Assuming the data is:
Gender <- c("M", "F", "F", "M", "F", "M", "F", "M")
Shoe_Size <- c(10, 7, 6, 11, 7, 12, 8, 10)
Height <- c(178, 160, 165, 185, 168, 190, 170, 180)

# Create the data frame
household_data <- data.frame(Gender, Shoe_Size, Height)
print(household_data)
```

#.1.cFind the mean of shoe size and height of the respondents. Write the R scripts and its result.
```{r}
# Subset for Males
males_subset <- subset(household_data, Gender == "M")
print(males_subset)

# Subset for Females
females_subset <- subset(household_data, Gender == "F")
print(females_subset)
```

#d. Is there a relationship between shoe size and height? Why?
```{r}
# Calculate the mean of Shoe_Size
mean_shoe_size <- mean(household_data$Shoe_Size)

# Calculate the mean of Height
mean_height <- mean(household_data$Height)

# Print the results
cat("Mean Shoe Size:", mean_shoe_size, "\n")
cat("Mean Height:", mean_height, "\n")
```

#.2
```{r}
# Original character vector
months_vector <- c("March","April","January","November","January",
                   "September","October","September","November","August",
                   "January","November","November","February","May","August",
                   "July","December","August","August","September","November","February",
                   "April")

# Convert to a factor
factor_months_vector <- factor(months_vector)

# Print the factor vector
print(factor_months_vector)
```


#3
```{r}
# Summary of the character vector
summary_char <- summary(months_vector)
print(summary_char)

# Summary of the factor vector
summary_factor <- summary(factor_months_vector)
print(summary_factor)
```

#4
```{r}
# Create the character vector with repeated directions based on frequency
factor_data <- c("East", 
                 "West", "West", "West", "West",
                 "North", "North", "North")

# Apply the factor function with the required order of the level
new_order_data <- factor(factor_data, levels = c("East","West","North"))

# Print the result
print(new_order_data)
```

#5.a
```{r}
#To import a CSV file (Comma Separated Values) using read.table():
#read.table() is a general function. For CSV, we set sep="," and header=TRUE.
#If the file were tab-separated, we would use sep="\t".
data_imported <- read.table("import_march.csv", header = TRUE, sep = ",")
```
#5.b
```{r}
# R script to view the dataset
View(data_imported)
# or, for a console print
head(data_imported)
```

#6
```{r}
exhaustive_search <- function(user_input) {
  # 1. Check if the number is beyond the range of 1 to 50
  if (user_input < 1 || user_input > 50) {
    print("The number selected is beyond the range of 1 to 50")
  } 
  # 2. Check for the specific number 20
  else if (user_input == 20) {
    print("TRUE")
  } 
  # 3. Otherwise (within range 1-50 but not 20), display the input number
  else {
    print(user_input)
  }
}

# Example 1: Randomly select a number (from 1 to 50) and display it
random_num <- sample(1:50, 1)
cat("The chosen number is:", random_num, "\n")
exhaustive_search(random_num)

# Example 2: Input is 20
cat("\nInputting 20:\n")
exhaustive_search(20)

# Example 3: Input is beyond the range
cat("\nInputting 60:\n")
exhaustive_search(60)
```

